---
title: "SCEM Final Coursework"
author: "Sangbin Lim (2747409)"
subtitle: Task II
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

*****

## IMPORTANT NOTES: 
- **DO NOT** change the code block names. Enter your solutions to each question into the predefined code blocks. 
- **Don't forget to replace the placeholder value in code block "Setup" by your valid student ID number.**
- If you want to use any packages, add their names to the vector `required.packages` in the code block "Setup". That code block will 
automatically install those packages (if needed) and load them. Any other calls to `install.packages()`, `library()` or `require()` may result in zero marks for the activity where it happens.
- Make sure that the data files *df.rds* and *df_holdout.rds* are located in the same folder as this file, _Task02.Rmd_.
- Instructions for this task are provided in blue text below. Your solution should NOT be added within those text boxes, and should NOT be in blue text. **During marking, all the text within the text blocks below, with names starting with "prompt", will be removed and not visible to the marker.**

```{r ID, echo=FALSE, message=FALSE, warning=FALSE}
## We will use your student ID as the seed for the random number generator.
MY_STUDENT_ID <- 2747409 # <-- Replace "ABCDEFG" by your student ID number (as an integer, i.e., without quotes).
```

```{r Setup, echo=FALSE, message=FALSE, warning=FALSE}
force_reinstall_everything <- FALSE ## <--- change to TRUE to force a reinstall of 
                                    ## all packages listed in allowed.packages

## ADD ANY REQUIRED PACKAGES TO THE VECTOR BELOW
## Note: tidyverse includes:  ggplot2, dplyr, tidyr, readr, purrr, tibble, 
##                            stringr, forcats, and lubridate.
allowed.packages <- c("tidyverse", 
                      "tidymodels", 
                      "recipes", 
                      "parsnip", 
                      "bonsai", 
                      "healthyR.ai", 
                      "knitr",
                      "DataExplorer",
                      "visdat"
                      )

## DO NOT CHANGE ANYTHING ELSE IN THIS CODE BLOCK
for (i in seq_along(allowed.packages)){
  if(!(allowed.packages[i] %in% installed.packages()[, 1]) || force_reinstall_everything){
    install.packages(allowed.packages[i], 
                     dependencies = c("Depends", "Imports"),
                     repos = "https://cloud.r-project.org")
  }
  library(allowed.packages[i], character.only = TRUE)
}

knitr::opts_chunk$set(echo = TRUE, collapse=TRUE)
ignore <- runif(1) # To initialise the PRNG

cat("\nList of allowed packages (pre-loaded):\n ", paste(allowed.packages, collapse="\n  "))
```

```{r checkID, echo=FALSE}
if(!is.numeric(MY_STUDENT_ID)){
  stop("MY_STUDENT_ID MUST BE YOUR VALID NUMERIC ID.")
} 
```

*****

:::{#prompt .message style="color: blue;"}
In this task you will build a full predictive pipeline for the problem described in the coursework specs. Please read each activity carefully and follow the instructions to complete your coursework.

Make sure that the data file **df.rds** is placed in the same folder as this Task02.Rmd file. The data that you can use for model development in this task will then be automatically loaded and set up for you, and stored into a dataframe variable called `df`.
:::


```{r load_data, echo = FALSE}
## DO NOT CHANGE ANYTHING IN THIS CODE BLOCK
set.seed(MY_STUDENT_ID)
df <- readRDS("df.rds") %>% sample_frac(0.6, replace = FALSE)
```

*****

## 1. Exploratory Data Analysis
:::{#prompt .message style="color: blue;"}
**Note**: It is expected that this section should result in no more than 600-ish words (possibly less; not counting the code blocks).

Add here your exploratory data analysis. Your solution should include both your code (which markers must be able to run independently) and your rationale for each exploratory step. In general, your EDA steps should include:

- A short justification of what a given plot or statistical summary is intended to reveal about the data.

- A code block implementing your EDA step.

At the end of your EDA section, you should also include a short commentary on what your data exploration revealed about the data, and which pre-processing steps may be required based on that.

You can use as many code blocks as needed, but don't overdo it (remember - everything needs to have a proper rationale). 
:::

### 1.1 Data overview

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The dataset is derived from a binary epitope prediction problem, where each observation is described by a large set of numerical predictive features. An exploratory data analysis
provided an initial overview of potential dependencies caused by "Info_group", distribution of the target "Class", and the behavior of features "feat_" in terms of scale, variability and missingness.
:::

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Quick overview shows that some of the columns that begin with "Info_" are mainly identifiers or meta data. As these variables do not carry predictive information about the target data, discarding them could be considered. Moreover, further investigation is required as the presence of some missing values and variables with very different scales could be identified from the dataset. Data preprocessing such as imputation and standardisation can be used to ensure that the predictors are on comparable scales and that missing values do not bias the prediction model.
:::

```{r, echo = FALSE}

glimpse(df)

```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The dataset is high-dimensional with 7403 rows and 390 columns including 2 discrete and 388 continuous variables. Since most of the features are numerical, no categorical encoding has to be done. Instead, preprocessing can focus on handling missing values, standardising feature scales, and applying dimensionality reduction.
:::

```{r, echo = FALSE}

df_summary <- introduce(df) %>% tibble()
t(df_summary)

```

### 1.2 Class distribution and grouping structure

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Analysis of distribution of the target variable "Class" and summary of the grouping variable "Info_group" were conducted to assess the target distribution proportion and existence of dependencies within the dataset.
:::

```{r, echo = FALSE}
# Class distribution
df %>% count(Class, name = "Count") %>% mutate(Distribution_Proportions = percent((Count/sum(Count))))

```

```{r, echo = FALSE}
# Summary stats of "Info_group" 
df %>% count(Info_group) %>% summary()

```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The distribution of the variable "Class" seems fairly balanced, showing a split of 59% for the negative and 41% for the positive class. In contrast, the summary of "Info_group" provides clear evidence of dependencies within the dataset, with the number of observations ranging from 1 to 632, with an average of approximately 13.53 observations per group. Therefore, the assumption of independence among the observations is likely to be violated if each observation is treated independently. Hence, to avoid data leakage, splitting, preprocessing and modelling should be done at the group level so that all observations from the same group are assigned either to the training or to the evaluation set.
:::

### 1.3 Missing values

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Analysis of the missing values were conducted to assess the amount and pattern of missing values in the dataset.
:::

```{r, echo = FALSE}

# Total Missing rate
total_na_rate <- df_summary$total_missing_values  / df_summary$total_observations
percent(total_na_rate, accuracy = 0.01)

```

```{r, echo = FALSE}

# Each variable's missing rate ratio
variable_na_rate <- tibble(
  Variable    = names(df),
  NA_ratio = percent(sapply(df, \(x) mean(is.na(x)) * 100), accuracy = 0.01)
)

variable_na_rate

```

```{r, echo = FALSE}

# Top 20 of variable missing rate
top_na_rate <- variable_na_rate %>%
  filter(NA_ratio > 0) %>%
  arrange(desc(NA_ratio)) %>%
  slice_head(n = 20) %>%
  mutate(Variable = fct_reorder(Variable, NA_ratio))

ggplot(top_na_rate, aes(x = Variable, y = NA_ratio)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  ylab("Missing Rate") +
  xlab("Variable") +
  theme(axis.text.y = element_text(hjust = 1))

```


```{r, echo = FALSE}

# Rows with NA in target and non-feature variables
colSums(is.na(df[, c("Info_PepID", "Info_pos","Info_protein_id", "Info_group", "Class")]))
df %>% filter(is.na(Info_pos)|is.na(Info_protein_id))

```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The overall proportion of missing values ($\approx$ 0.02%) is extremely low with 499 missing values out of 2,887,170 total entries. Each variable's missing rates tend to be very low and are mostly found in a subset of feature variables. This suggests that a simple imputation strategy, such as median imputation for numerical features, should be sufficient to handle the few missing values, without needing to drop them. Furthermore, no concerning missing values were found in the target and non-feature variables. Therefore, since missingness is limited to a few feature variables only, the next important aspect to examine is the distribution and scale of the features for feature scaling and dimensionality reduction.
:::

### 1.4 Feature distributions and scale

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Analysis of the distributions and scales of the feature variables was conducted to assess their ranges, variability and differences in the scale.
:::

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Summary statistics of features including minimum, maximum, mean, and range were computed to identify the characteristics of each feature's scale. 
:::

```{r, echo = FALSE}

# Summary stats of feature variables
feat_variables <- df %>% select(starts_with("feat_"))

feat_stats  <-
  tibble(
    Variable  = names(feat_variables),        
    Min  = apply(feat_variables, 2, min, na.rm = TRUE),  
    Max  = apply(feat_variables, 2, max, na.rm = TRUE),
    Mean = apply(feat_variables, 2, mean, na.rm = TRUE),
    Range = Max - Min
  ) 

feat_stats
```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The histogram of feature ranges showed variations of their range which indicated heterogeneity in scale across features.
:::

```{r, echo = FALSE}

# The histogram of feature range
ggplot(feat_stats, aes(x = Range)) +
  geom_histogram(bins = 50) +
  scale_x_log10() +
  theme_bw() +
  labs(x = "Feature range (log10 scale)", y = "Count")

```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
To take a closer look into the features scale, top 20 features with the largest range were extracted and plotted. The result showed highlighted features such as "feat_local_SOCN_Grantham.lag" and "feat_local_MolWeight" that could lead to information overload on the model due to wide spectrum of its range compared to other features.    
:::

```{r, echo = FALSE}

# Top 20 of feature range
top_feat_range <- feat_stats %>%
  arrange(desc(Range)) %>%
  slice_head(n = 20) %>%
  mutate(Variable = fct_reorder(Variable, Range))

top_feat_range

```

```{r, echo = FALSE}

# Top 20 of feature range plot
ggplot(top_feat_range, aes(x = Variable, y = Range)) +
  geom_col() +
  coord_flip() +
  scale_y_log10() +
  theme_bw() +
  labs(
    title = "Top 20 of feature range (log10 scale)",
    x = "Feature",
    y = "Range"
  )

```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
To check whether range difference is due to outliers, ranges were contrasted with truncated ranges. Removal of the upper/lower 0.1% of the data lead to the range reduction concentrated between 10 to 30%, indicating that most of the features show small change in scale when their tails are trimmed, while some show large reduction due to influence of their outliers.
:::

```{r, echo = FALSE}

# Function for trimmed range
myrange <- function(x, trim = 0){
  x <- x[!is.na(x)]          # NA 제거
  n  <- length(x)
  lo <- floor(n * trim) + 1
  hi <- n + 1 - lo

  return(diff(range(sort.int(x)[lo:hi])))
}

# Range comparison between feature variables
tmp_range <- tibble(
  Variable  = names(feat_variables),
  Range_00  = apply(feat_variables, 2, myrange, trim = 0),
  Range_001 = apply(feat_variables, 2, myrange, trim = 0.001)) %>%
  mutate(RangeRed_percent = (1 - Range_001 / Range_00) * 100)

# Range reduction after truncating 0.01% upper/lower tails histogram plot
ggplot(tmp_range, aes(x = RangeRed_percent)) +
  geom_histogram(bins = 50) +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
  labs(title = "Range reduction after truncating 0.1% upper/lower tails", x = "Range reduction (%)", y = "Count")


```

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The overview of feature distributions and scales show heterogeneous scales and high dimensionality. Hence, feature standardisation and dimensionality reduction will be performed to control the scale difference, redundancy and stability of the model.
:::

### 1.5 EDA summary and implications for preprocessing

:::{#Exploratory_Data_Analysis_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Based on the EDA, the following aspects of the dataset should be taken into account during preprocessing:

* Discard "Info_" variables that act primarily as identifiers / meta data.
* Split train and test data at the group level of "Info_group" to avoid data leakage.
* Apply simple median imputation for missing features.
* Standardise features to a common scale and conduct a dimensionality reduction such as PCA to reduce redundancy and increase stability to the model.

:::

*****

## 2. Data preprocessing and feature engineering

:::{#prompt .message style="color: blue;"}
**Note**: It is expected that this section should result in no more than 400-ish words (possibly less; not counting the code blocks). 

Add here your data pre-processing and feature selection/engineering steps. Your solution should include both your code (which markers must be able to run independently) and your rationale. In general, your steps should include:

- A short justification of which insight from your EDA is motivating a given data transformation.

- A code block implementing your pre-processing step.

**In addition** to the regular data transformations emerging from the insights you gathered during EDA, this section must contain the splitting off of a test set that you'll use later for final performance assessment  (check the CW specs for details).

At the end of this section you should also include a short commentary on what changes the pre-processing has induced in your data (i.e., what did the data "look like" before this step vs. what it "looks like" afterwards.

You can use as many code blocks as needed, but don't overdo it (remember - everything needs to have a proper rationale). If you compare different pre-processing options, make sure to make it clear, and to indicate at the end which option you selected, and why. 

**Note**: In this section you'll be basically defining which data transformations you will later incorporate into your predictive pipeline. When you build the pipeline in the next section, you'll encapsulate all those steps (together with the model) into a single object that you can fit and deploy as a single unit.

### 2.1 Split train and test data at the group level

:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Since a grouping structure exists in the dataset, the data is split into training and test sets at the group level before the sequential data preprocessing to minimise data leakage. However, the preprocessing in this section follow the predefined data structure and assumptions in the task specification. In particular, the target Class is converted into a factor and identifier columns that are not informative are excluded, leaving only numerical features to be used as inputs for modelling.
:::

```{r, echo = FALSE}
# Predefined data structure and assumptions.
# 1) Two ID columns named “Info_PepID” and “Info_pos”. Together, these two columns provide a unique reference to each row.
# 2) One Group column named “Info_group”
# 3) Other columns starting with “Info_” that are not informative and should be discarded.
# 4) A few hundred feature columns with names starting with “feat_”, containing numerical predictive features that were calculated for each observation.
# 5) 1 Class column, containing the target class.

non_inform_var <- c("Info_PepID", "Info_pos", "Info_protein_id")
df_raw <- df

# Mutate unique ID column, drop non informative variables and convert "Class" type into a factor.
df_modified <- df_raw %>% 
  mutate(
    Info_ID = paste(Info_PepID, Info_pos, sep = "_"),
    Class = factor(Class, 
                   levels = c(1, -1),
                   labels = c("epitope", "non-epitope")), 
    Info_group = as.factor(Info_group)) %>%
  relocate(Info_ID) %>%
  select(-all_of(non_inform_var))

glimpse(df_modified[, c("Info_ID", "Info_group", "Class")])
```


```{r, echo = FALSE}
set.seed(MY_STUDENT_ID)

split <- group_initial_split(df_modified, group = Info_group, prop = 0.8)

df.train <- training(split)
df.test <- testing(split)

```

:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
This split ensures that no Info_group appears in both training and test sets, preserving the independence of the final evaluation. From now onward, only the training data "df.train" will be used for data preprocessing and feature engineering.
:::

### 2.2 Preprocessing recipe

:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Based on the EDA conducted, a recipe is created to build preprocessing workflow for data. The recipe selects the "feat_" columns as predictors, imputes missing values, standardise the predictors scales and conduct PCA for dimensionality reduction. For imputation, since only a small portion of missing values and outliers exist, median imputation was chosen as a more robust option than the mean, because it is less sensitive to extreme values. Moreover, although trimming or winsorisation to control outlier was considered, since their wide ranges seem to be natural due to their characteristics, it was not applied. Instead, standardisation and PCA was conducted.
:::

```{r, echo = FALSE}
# Initial Recipe to scrutinize the data for modelling.  
initial_recipe <- recipe(Class ~ ., data = df.train) %>%
  # Exclude informative data from predictors
  update_role(Info_ID, new_role = "id") %>%
  update_role(Info_group, new_role = "group") %>%
  # Impute missing values through median imputation.
  step_impute_median(all_numeric_predictors()) %>%
  # Standardise predictors scales.
  step_normalize(all_numeric_predictors()) %>%
  # Retain predictors that have >=95% of total variance.
  step_pca(all_numeric_predictors(), threshold = 0.95)

initial_recipe 
```

```{r, echo = FALSE}
# Initial Prep
initial_recipe_prepped <- initial_recipe %>% prep()

summary(initial_recipe_prepped)
```

```{r, echo = FALSE}
# To confirm normality, predictor's mean should be approximately 0 and standard deviations at 1.
test_recipe_without_pca <- recipe(Class ~ ., data = df.train) %>%
  update_role(Info_ID, new_role = "id") %>%
  update_role(Info_group, new_role = "group") %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

test_recipe_without_pca_prep   <- prep(test_recipe_without_pca)
test_recipe_without_pca_bake <- bake(test_recipe_without_pca_prep, df.train)

test_recipe_without_pca_bake %>%
  select(starts_with("feat_")) %>%
  summarise(across(everything(), list(mean = mean, sd = sd))) %>%
  t() %>%
  head(10)

```
:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Further inspection on PCA's cumulative variance is conducted to identify the degree of contribution of each variable and ideal number of principal components (PCs) to obtain roughly 90 to 95% of the information.
:::

```{r, echo = FALSE}
# Find where step for PCA is located in the recipe.
pca_step <- which(sapply(initial_recipe_prepped$steps, inherits, "step_pca"))

# Variance explained by each PC and its cumulative variance.
pca_variance <- initial_recipe_prepped$steps[[pca_step]]$res
summary(pca_variance)

```

```{r, echo = FALSE}
# The degree of contribution of each variable.
screeplot(pca_variance, 
          main = "Scree plot of PCA",
          col = "steelblue", 
          type = "lines", 
          pch = 1, 
          npcs = length(pca_variance$sdev),
          )
```

```{r, echo = FALSE}
# Location of PCs explaining 90 and 95 of the variance.
pca_importance <- summary(pca_variance)$importance
cum_variance <- pca_importance["Cumulative Proportion", ]
cum_variance_90 <- which(cum_variance >= 0.90)[1] # PC96
cum_variance_95 <- which(cum_variance >= 0.95)[1] # PC130

pca_importance[, c(cum_variance_90, cum_variance_95)]
```
:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Summary of the PCs and its scree plot indicate that roughly one third of the PCs account for most of the variability in the predictors. PC130’s standard deviation tells us it’s 0.63 dispersed in relation to the mean of the data, proportion of variance tells us the influence contributed and cumulative proportion tells us that 95% of the data is explained by PC130.
:::

:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
To wrap up, perform a quick data inspection to verify whether the initial recipe is properly implemented. 
:::

```{r, echo = FALSE}
# Initial Bake 
initial_recipe_baked <- bake(initial_recipe_prepped, df.train)
summary(initial_recipe_baked)

```

```{r, echo = FALSE}
# Check for the data structure and properties.
str(initial_recipe_baked)
head(initial_recipe_baked)
```

```{r, echo = FALSE}
# Check for the missing value.
sum(is.na(initial_recipe_baked))
```

:::{#Data_preprocessing_and_feature_engineering_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Overall, the recipe function transforms the original high-dimensional dataset into a standardised version ideal for modelling. Now, with the analysis done in data preprocessing, pipelines using PCA thresholds of approximately 90% and 95% cumulative variance will be compared with grouped cross-validation and F1-score then the configuration with better performance will be selected as the final model.
:::

*****

## 3. Modelling

:::{#prompt .message style="color: blue;"}
**Note**: It is expected that this section should result in no more than 400-ish words (possibly less; not counting the code blocks).

Add here your modelling steps. Your solution should include both your code (which markers must be able to run independently) and your rationale. This section must include:

- The fitting of a preliminary model on your preprocessed training data (to estimate some baseline performance level and to make sure your data is ready for modelling). Make this a simple model - logistic regression, decision trees or kNN are good options (but other may be used instead, if you prefer).

- The building of a predictive pipeline encapsulating all the learned/learnable aspects of your solution. This should result in a pipeline-type object that uses your raw training data (post EDA and test set splitting, but prior to any preprocessing), fits all your preprocessing transformations (incl. dimensionality reduction) and the predictive model using that data, and can then be used to ingest your test data (or any new data in the same format) and generate predictions for each entry. 

Please refer to the CW specs for other requirements of this step (including, e.g., the addition of a dimensionality reduction step into your pipeline).

You can use as many code blocks as needed, but don't overdo it (remember - everything needs to have a proper rationale). If you compare different models, make sure to make it clear, and to indicate at the end which pipeline option is your final selected one.

:::

### 3.1 Grouped cross-validation

:::{#Modelling_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The performance of the models are assessed using grouped 5-fold cross-validation with F1 score, accuracy, sensitivity, specificity and ROC curve as a metric.
:::

```{r, echo = FALSE}
set.seed(MY_STUDENT_ID)

# Grouped 5-fold cross validation
folds <- group_vfold_cv(df.train, 
                           group = Info_group, 
                           v = 5)
```

### 3.2 Baseline model: Logistic regression

:::{#Modelling_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Before the finalisation of the model, with Logistic regression as the baseline model, PCA retaining 90% and 95% of the variance is compared. Besides showing the ideal number of PCs required for the model, these results also serve as a validation that the preprocessing recipe is working as intended.
:::

```{r, echo = FALSE}
cum_variance_90_recipe <- test_recipe_without_pca %>%
  step_pca(all_numeric_predictors(), num_comp = cum_variance_90)

cum_variance_95_recipe <- test_recipe_without_pca %>%
  step_pca(all_numeric_predictors(), num_comp = cum_variance_95)

# Logistic regression model set up
log_reg_model <- logistic_reg(mode = "classification") %>% set_engine("glm")

cum_variance_90_wf <- workflow() %>%
  add_model(log_reg_model) %>%
  add_recipe(cum_variance_90_recipe)

cum_variance_95_wf <- workflow() %>%
  add_model(log_reg_model) %>%
  add_recipe(cum_variance_95_recipe)

```

```{r, echo = FALSE}
# Fit the model using cross-validation.
set.seed(MY_STUDENT_ID)

metric_assess <- metric_set(accuracy, roc_auc, ppv, npv, f_meas)

cum_variance_90_fit <- cum_variance_90_wf %>%
  fit_resamples(folds, metrics = metric_assess)

cum_variance_95_fit <- cum_variance_95_wf %>%
  fit_resamples(folds, metrics = metric_assess)

```


```{r, echo = FALSE}
# Fit metrics of PCA retaining 90% of the variance.
collect_metrics(cum_variance_90_fit)
```

```{r, echo = FALSE}
# Fit metrics of PCA retaining 95% of the variance.
collect_metrics(cum_variance_95_fit)
```

```{r, echo = FALSE}
# PCA retaining 90% of the variance shows the ideal number of PCs required.
log_reg_model_90_wf <- cum_variance_90_wf
log_reg_model_90_fit <- cum_variance_90_fit 
log_reg_model_90_metrics <- collect_metrics(cum_variance_90_fit)
```

:::{#Modelling_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Since grouped 5-fold Cross-validation show almost identical accuracy, F1 and ROC AUC for the 90% and 95% PCA thresholds, the lower-dimensional 90% model will be chosen for the final recipe.
:::

### 3.3 Model comparison: logistic vs random forest 

:::{#Modelling_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
To assess whether predictive performance varies across different modelling approaches, a penalized logistic regression model and random forest model were compared against a baseline logistic regression model.
:::

```{r, echo = FALSE}
# Final Recipe to be used.
final_recipe <- recipe(Class ~ ., data = df.train) %>%
  update_role(Info_ID, new_role = "id") %>%
  update_role(Info_group, new_role = "group") %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_pca(all_numeric_predictors(), num_comp = cum_variance_90)

```

```{r, echo = FALSE}
# Logistic regression model set up.
final_log_reg_model <- logistic_reg(mode = "classification") %>% set_engine("glm")

final_log_reg_model_wf <- workflow() %>%
  add_model(final_log_reg_model) %>%
  add_recipe(final_recipe)

```

```{r, echo = FALSE}
# Fit the Logistic regression model using cross-validation.
set.seed(MY_STUDENT_ID)

final_log_reg_model_fit <- final_log_reg_model_wf %>%
  tune_grid(resamples = folds,
            metrics = metric_assess)

```

```{r, echo = FALSE}
collect_metrics(final_log_reg_model_fit)
```

```{r, echo = FALSE}
# penalized logistic regression model set up.
pen_log_model <- logistic_reg(
  mode    = "classification",
  penalty = tune(), 
  mixture = 0         
) %>%
  set_engine("glmnet")

pen_log_model_wf <- workflow() %>%
  add_model(pen_log_model) %>%
  add_recipe(final_recipe)

# set penalty parameter as -4 to 0.
pen_log_grid <- grid_regular(
  penalty(range = c(-4, 0)),
  levels = 20
)

```

```{r, echo = FALSE}
# Fit the penalized logistic regression model using cross-validation.
set.seed(MY_STUDENT_ID)

pen_log_model_fit <- pen_log_model_wf %>% 
  tune_grid(resamples = folds, 
            grid = pen_log_grid, 
            metrics = metric_assess)
```

```{r, echo = FALSE}
collect_metrics(pen_log_model_fit)
```

```{r, echo = FALSE}
# Random forest model set up.
rf_model <- rand_forest(mtry = tune(),
                             min_n = tune(),
                             trees = 100,
                             mode = "classification") %>% set_engine("ranger")

rf_model_wf <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(final_recipe)

# set mtry as sqrt(96), thus 10 up to 60 and min_n 5 to 20 to balance over/under-fitting.
rf_grid <- grid_regular(
  mtry(range = c(10, 60)),
  min_n(range = c(5, 20)),
  levels = 2
)
```

```{r, echo = FALSE}
# Fit the random forest model using cross-validation.
set.seed(MY_STUDENT_ID)

rf_model_fit <- rf_model_wf %>% 
  tune_grid(resamples = folds, 
            grid = rf_grid, 
            metrics = metric_assess)
```

```{r, echo = FALSE}
collect_metrics(rf_model_fit)
```

```{r, echo = FALSE}
# Logistic Regression model's best F1
log_reg_best_f1 <- collect_metrics(final_log_reg_model_fit) %>%
  filter(.metric == "f_meas") %>%
  arrange(desc(mean)) %>%
  slice(1) %>%
  mutate(model = "logistic")

# Penalized Logistic Regression model's best F1
pen_log_best_f1 <- collect_metrics(pen_log_model_fit) %>% 
  filter(.metric == "f_meas") %>%
  arrange(desc(mean)) %>%
  slice(1) %>%
  mutate(model = "penalized_logistic")

# Random Forest model's best F1
rf_best_f1 <- collect_metrics(rf_model_fit) %>%
  filter(.metric == "f_meas") %>%
  arrange(desc(mean)) %>%
  slice(1) %>%
  mutate(model = "random_forest")

model_comparison <- bind_rows(log_reg_best_f1, pen_log_best_f1, rf_best_f1)
model_comparison
```

:::{#Modelling_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The comparison of F1-score between the two models show that the random forest model with F1 score $\approx$ 0.60 has better score than logistic regression model with $\approx$ 0.48 and penalized logistic regression model with $\approx$ 0.47 . Therefore, the random forest model is selected for further evaluation.
:::

*****

## 4. Model assessment

:::{#prompt .message style="color: blue;"}
**Note**: It is expected that this section should result in no more than 100-ish words (possibly less; not counting the code blocks).

This is a short section in which you must estimate the generalisation performance of your model on the test set that you isolated at the start of your Data preprocessing step. It is also a chance for you to ensure that your pipeline object can take in new data and return the required predictions.

Use your pipeline to generate predictions for your test set, and calculate the observed performance using the $F_1$-score as your main performance indicator. You can also additionally report other performance metrics if you think it's relevant.

You can use as many code blocks as needed, but it's likely that this section can be completed with a single one. Make sure that your performance indicators are clearly echoed, so that a reader can easily check if after re-running your script. Add a short comment on the observed performance (is it good? Is it poor? Remember that you're not being assessed on the performance of your models, but on the soundness of your approach).
:::

:::{#Model_assessment_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
Hyperparameter tuned random forest model is refitted on the whole training set with finalized preprocessing recipe. It is then used to generate predictions for the test data which will provide an unbiased estimate of how well the model is expected to perform on new, unseen data such as the held-out test set.
:::

```{r, echo = FALSE}
# Redefine the workflow and fit test using the optimised hyperparameters.
final_wf <- rf_model_wf %>% finalize_workflow(rf_best_f1)
final_fit <- final_wf %>% fit(data = df.train)
```


```{r, echo = FALSE}
set.seed(MY_STUDENT_ID)

wf_final_fit <- final_wf %>% last_fit(split, metrics = metric_assess)
wf_final_fit %>% collect_metrics()

# The estimated ROC curve for the final model
wf_final_fit %>% 
  collect_predictions() %>% 
  roc_curve(Class, .pred_epitope) %>% 
  autoplot()

```

```{r, echo = FALSE}
# Prediction
test_predictions <- predict(final_fit, df.test, type = "prob") %>%
  bind_cols(predict(final_fit, df.test, type = "class")) %>%
  bind_cols(df.test %>% select(Class))

test_results <- metric_assess(
  data    = test_predictions,
  truth   = Class,
  estimate = .pred_class,
  .pred_epitope)

test_results
```


```{r, echo = FALSE}
# Confusion matrix
test_conf_mat <- conf_mat(test_predictions, truth = Class, estimate = .pred_class)
test_conf_mat
```

:::{#Model_assessment_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The final random forest model achieved an accuracy of 0.66, F1-score of $\approx$ 0.60, precision(PPV) of $\approx$ 0.69 and Negative predictive value(NPV) of $\approx$ 0.65, with a ROC AUC of $\approx$ 0.76. The confusion matrix shows that the model correctly identifies about half of the true epitopes (true positives vs false negatives) and reasonably often rejects non-epitopes (true negatives vs false positives), indicating moderate rather than outstanding predictive performance. This moderate performance may be due to difficulty of the epitope prediction problem, the limited discriminative power of the available features, and the relatively simple tuning of the random forest hyperparameters. Overall, these results suggest that the model captures some useful signal for epitope prediction but still leaves substantial room for improvement.
:::

*****

## 5. Model deployment

:::{#prompt .message style="color: blue;"}
**Note**: This section is not expected to have much text in it, only the code and the results.

Finally, in this section you'll generate predictions for some data for which you do not know the class labels. This simulates the real-life scenario in which your model needs to be deployed to generate new predictions for new, unknown data (after all, if the labels were known, we wouldn't need a model).

Make sure that the data file __df_holdout.rds__ is placed in the same folder as this Task02.Rmd file. The new data will then be automatically loaded and set up for you, and stored into a dataframe variable called `df_holdout`, with the same columns as the original `df` that was loaded at the start of this task.

- Use your fitted pipeline to generate predictions for this new data.

- Build a new data frame called containing the following columns:
    - `Info_PepID` (taken from df_holdout)
    - `Info_pos`  (taken from df_holdout)
    - `Predicted_class` (with the predictions produced by your pipeline)

- Save your predictions dataframe to a file called __mypreds.rds__. You will be able to submit this file, and it will be used to give you some feedback on the actual performance of your pipeline on new data.
:::

:::{#Model_deployment_txt .message style="color: black; border: 1px outset black; background-color: #eeeffe;"}
The last step is to deploy the final trained model to the unseen holdout data ("df_holdout") to generate predictions, which are then saved for external evaluation.
:::

```{r load_holdout, echo = FALSE}
## DO NOT CHANGE ANYTHING IN THIS CODE BLOCK
df_holdout <- readRDS("df_holdout.rds")
head(df_holdout)
```


```{r, echo = FALSE}

df_holdout_modified <- df_holdout %>%
  mutate(
    Info_ID = paste(Info_PepID, Info_pos, sep = "_"),
    Info_group = as.factor(Info_group)) %>%
  relocate(Info_ID) %>%
  select(-all_of(non_inform_var))

glimpse(df_holdout_modified[, c("Info_ID", "Info_group")])
```
```{r, echo = FALSE}
# Prediction.
df_holdout_predictions <- predict(final_fit, df_holdout_modified, type = "class")

# The new data frame to be submitted.
df_holdout_result <- df_holdout %>%
  select(Info_PepID, Info_pos) %>%
  mutate(Predicted_class = df_holdout_predictions$.pred_class)

df_holdout_result


```

```{r, echo = FALSE}
# Save the result to RDS file.
saveRDS(df_holdout_result, file = "mypreds.rds")

```

*****
*****

### =========== End of Task II ===========
